---
title: "data summary"
author: "Stephanie Copeland"
date: "2023-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Packages
```{r}
library(here)
library(tidyverse)
library(ggplot2)
```
##Read in csv file
###information
*put information on csv file here*
```{r}
sampler_mos_counts <- read_csv(
  here::here("Data/sampler_mos_counts.csv"))
```

#### look at characteristics of current data set

```{r}
class(sampler_mos_counts)
#dataframe
head(sampler_mos_counts)
```

sampler - character: Initial ID of which researcher was sampling
SC = Steph Copeland (project lead)
MR = Matthew Rosen (undergraduate techician)
SL = Sylvia Li (undergraduate technician)
ST = Sachi Takishima (volunteer)
CC = Charlie Copeland (volunteer, Steph's dad)
KK = Kayla Kauffman (volunteer)
also present for all sampling endeavors but not a sampler was Hugo, Steph's service dog. He is a 70lbs German Shepherd Mix 
Lake - character: name of lake which collections were taken. Heart, Hidden, Flower, No_Name, Gilbert, Matlock, Marsh, LEB (Lower Eastern Brook), Little_Fish, Little_Lakes, Mack, EB (Eastern Brook)
Date - character: changed to date class (see below) date at which sampling occurred
Time_start - time: time that sampling began
Time_end = time: time that sampling commenced
Total_time - numeric: the amount of minutes sampled, calculated by determining the number of minutes that had passed between Time_start and Time_end
snow_melt_mosq - numeric: count of caught mosquitos falling into this phenotypic category
black_mosq - numeric: count of caught mosquitos falling into this phenotypic category
culiseta - numneric: count of mosquitos falling into this genera 
culex_tarsalis - numeric: count of caught mosquitos falling into this species 
blood_test - numeric: number of mosquitos removed from each sampling effort to test for the content of a blood meal. These samples were inherently destroyed by puncturing the abdomen of the sample with a burn/sterile pipette tip 

This is ok. I would like to clean column names and change the 'Date' column characterization from character to date

```{r, s/o to chatGPT for a large chunk of this code...}
sampler_mos_counts <- sampler_mos_counts %>%
  janitor::clean_names()
#now all lower case column names

sampler_mos_counts$date <- gsub("-", " ", sampler_mos_counts$date)
#date data was initially recorded as, for example, "4-Jul or 22-Aug", this removed the dash. 

addLeadingZeros <- function(x) {
  parts <- strsplit(x, " ")
  days <- sapply(parts, function(part) {
    if (nchar(part[1]) == 1) {
      sprintf("%02d", as.numeric(part[1]))
    } else {
      part[1]
    }
  })
  paste(days, parts[[1]][2])
}

sampler_mos_counts$date <- addLeadingZeros(sampler_mos_counts$date)
#created a function to add a 0 to any single digit date entry. For example an entry of "4-Jul" was changed to "04=Jul".

year <- 2023

sampler_mos_counts$date <- paste0(sampler_mos_counts$date, " ", year)
#added the year 2023 to each data entry in the date column. Column is still recognized as a character column. 

sampler_mos_counts$date <- as.Date(sampler_mos_counts$date, format = "%d %b %Y")
#changed date to a date column
```

#### Data set rearranging
Eventually, I would like to investigate the affects of sampler and time during day when sampled on mosquito counts. However, for right now. I would like to simplify the data to get some visual summary results. To do this I will create a new df with this data called 'summ_mos_counts'

