---
title: "data summary"
author: "Stephanie Copeland"
date: "2023-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Packages
```{r}
library(here)
library(tidyverse)
library(ggplot2)
library(lubridate)
```
##Read in csv file
###information
*put information on csv file here*
```{r}
sampler_mos_counts <- read_csv(
  here::here("Data/sampler_mos_counts.csv"))
```

#### look at characteristics of current data set

```{r}
class(sampler_mos_counts)
#dataframe
head(sampler_mos_counts)
```

- sampler - character: Initial ID of which researcher was sampling
- SC = Steph Copeland (project lead)
- MR = Matthew Rosen (undergraduate techician)
- SL = Sylvia Li (undergraduate technician)
- ST = Sachi Takishima (volunteer)
- CC = Charlie Copeland (volunteer, Steph's dad)
- KK = Kayla Kauffman (volunteer)
- also present for all sampling endeavors but not a sampler was Hugo, Steph's service dog. He is a 70lbs German Shepherd Mix 
- Lake - character: name of lake which collections were taken. Heart, Hidden, Flower, No_Name, Gilbert, Matlock, Marsh, LEB (Lower Eastern Brook), Little_Fish, Little_Lakes, Mack, EB (Eastern Brook)
- Date - character: changed to date class (see below) date at which sampling occurred
- Time_start - time: time that sampling began
- Time_end = time: time that sampling commenced
- Total_time - numeric: the amount of minutes sampled, calculated by determining the number of minutes that had passed between Time_start and Time_end
- snow_melt_mosq - numeric: count of caught mosquitos falling into this phenotypic category
- black_mosq - numeric: count of caught mosquitos falling into this phenotypic category
- culiseta - numneric: count of mosquitos falling into this genera 
- culex_tarsalis - numeric: count of caught mosquitos falling into this species 
- blood_test - numeric: number of mosquitos removed from each sampling effort to test for the content of a blood meal. These samples were inherently destroyed by puncturing the abdomen of the sample with a burn/sterile pipette tip 

This is ok. I would like to clean column names and change the 'Date' column characterization from character to date

```{r, s/o to chatGPT for a large chunk of this code...}
sampler_mos_counts <- sampler_mos_counts %>%
  janitor::clean_names()
#now all lower case column names

sampler_mos_counts$date <- gsub("(\\b\\d{1}\\b)", "0\\1", sampler_mos_counts$date)

sampler_mos_counts$date <- gsub("-", "", sampler_mos_counts$date)

sampler_mos_counts$date <- gsub("Jul", "07", sampler_mos_counts$date)
sampler_mos_counts$date <- gsub("Aug", "08", sampler_mos_counts$date)

sampler_mos_counts$date <- paste(sampler_mos_counts$date, "2023", sep = "")

sampler_mos_counts$date <- as.Date(sampler_mos_counts$date, format = "%d%m%Y")
```

#### Data set rearranging
Eventually, I would like to investigate the affects of sampler and time during day when sampled on mosquito counts. However, for right now. I would like to simplify the data to get some visual summary results. To do this I will create a new df with this data called 'summ_mos_counts'

```{r, I know this technically isnt in tidy format but I separate my code in a step-wise process. Other are welcome to combine this}
summ_mos_counts <- sampler_mos_counts %>% 
  select(-time_start, -time_end, -sampler)

summ_mos_counts <- summ_mos_counts %>% 
  group_by(lake, date) %>% 
  summarise(
    time_min = sum(total_time),
    total_snow_melt_mosq = sum(snow_melt_mosq),
    total_black_mosq = sum(black_mosq),
    total_culiseta = sum(culiseta, na.rm = TRUE),
    total_culex_tarsalis = sum(culex_tarsalis),
    total_blood_test = sum(blood_test, na.rm = TRUE)
  )

summ_mos_counts <- summ_mos_counts %>% 
  mutate(total_sum = total_snow_melt_mosq + 
           total_black_mosq +
           total_culiseta +
           total_culex_tarsalis +
           total_blood_test)

summ_mos_counts <- summ_mos_counts %>% 
  mutate(effort_min = total_sum/time_min)

summ_mos_counts$effort_min <- round(summ_mos_counts$effort_min, 1)
```
Describe resulting data set here 
added two additional columns and changed NAs to zeros in column total_blood_test
total_sum is the column adding all mosquitos across the phenotype groups and those destroyed for blood meal sampling through each row
effort/min is the column taking the total_sum column and dividing it by the time_min column for each row. This gets mosquitos caught per minute on the date the lakes were sampled

I actually also want to add a column to the df that notes 'fish' or 'no fish' 
```{r}
summ_mos_counts <- summ_mos_counts %>% 
  mutate(fish_status = case_when(
    lake %in% c("EB", "Hidden", "LEB", "Little_Lakes", "Matlock", "No_Name") ~ "no_fish",
    TRUE ~ "fish"
  ))
```

Average effort by lake
```{r}
avg_effort <- summ_mos_counts %>% 
  group_by(lake) %>% 
  summarise(avg_effort = mean(effort_min, na.rm = TRUE))
```

```{r}
ggplot(avg_effort, aes(x = lake, y = avg_effort))+
  geom_bar(stat = "identity", fill = "blue")+
  labs(title = "Effort by Lake", x = "Lake", y = "Effort (min)")+
  theme_minimal()
```

```{r}
avg_effort <- avg_effort %>% 
  mutate(fish_status = case_when(
    lake %in% c("EB", "Hidden", "LEB", "Little_Lakes", "Matlock", "No_Name") ~ "no_fish",
    TRUE ~ "fish"
  ))

ggplot(avg_effort, aes(x = lake, y = avg_effort, fill = fish_status))+
  geom_bar(stat = "identity")+
  labs(title = "Effort by Lake", x = "Lake", y = "Effort (min)")+
  theme_minimal()
```

standard error
```{r}
stan_error <- summ_mos_counts %>% 
  group_by(lake) %>% 
  summarize(standard_error = sd(effort_min) / sqrt (n()))
```

graph with error bars
```{r}
combo_avg_error <- merge(avg_effort, stan_error, by = "lake")

ggplot(combo_avg_error, aes(x = lake, y = avg_effort, fill = fish_status))+
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.6)+
  geom_errorbar(aes(ymin = avg_effort - standard_error, ymax = avg_effort + standard_error),
                position = position_dodge(width = 0.8),
                width = 0.3)+
  labs(title = "Average Effort by Lake", x = "Lake", y = "Average Effort")+
  theme_minimal()
```

graph with lakes in their "pairs"
```{r}
combo_avg_error$lake <- factor(combo_avg_error$lake, levels = c('EB', 'Mack', 'LEB', 'Marsh', 'Hidden', 'Heart', 'Little_Lakes', 'Little_Fish', 'Matlock', 'Flower', 'No_Name', 'Gilbert'))

ggplot(combo_avg_error, aes(x = lake, y = avg_effort, fill = fish_status))+
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.6)+
  geom_errorbar(aes(ymin = avg_effort - standard_error, ymax = avg_effort + standard_error),
                position = position_dodge(width = 0.8),
                width = 0.3)+
  labs(title = "Effort by Lake", x = "Lake", y = "Average Effort (mosquito/min)")+
  theme_minimal()
```

box plot - thanks Charlie for helping me come up with the ideas! 

```{r, each lake separate box plot}
ggplot(summ_mos_counts, aes(x = fish_status, y = effort_min, color = lake)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_point(position = position_dodge(width = 0.8), alpha = 0.7, size = 3) +
  labs(title = "Mosquito/Min by Fish Status", x = "Fish Status", y = "Effort (Mosquito/Min)") +
  theme_minimal()
```


```{r, two boxplots with position_jitterdodge, colored by lakes}
ggplot(summ_mos_counts, aes(x = fish_status, y = effort_min)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_point(aes(color = lake), position = position_jitterdodge(seed = 12), alpha = 0.7, size = 3) +
  labs(title = "Mosquito/Min by Fish Status", x = "Fish Status", y = "Effort (Mosquito/Min)") +
  theme_minimal()
```


```{r, two box plots with points colored by lakes}
ggplot(summ_mos_counts, aes(x = fish_status, y = effort_min)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_point(aes(color = lake), position = position_dodge(width = 0.8), alpha = 0.7, size = 3) +
  labs(title = "Mosquito/Min by Fish Status", x = "Fish Status", y = "Effort (Mosquito/Min)") +
  scale_color_manual(values = c("EB" = "blue", "Flower" = "#58593F", "Gilbert" = "red", "Heart" = "purple", "Mack" = "#155F83FF", "LEB" = "orange", "Marsh" = "magenta", "Hidden" = "pink", "Little_Lakes" = "gray", "Little_Fish" = "#4DAC26", "Matlock" = "#DFC27D", "No_Name" = "#D01C8B")) +
  theme_minimal()
```

```{r}
ggplot(summ_mos_counts, aes(x = fish_status, y = effort_min)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  geom_point(aes(color = lake), position = position_jitterdodge(seed = 12), alpha = 0.7, size = 3) +
  labs(title = "Mosquito/Min by Fish Status", x = "Fish Status", y = "Effort (Mosquito/Min)") +
  scale_color_manual(values = c("EB" = "blue", "Flower" = "#58593F", "Gilbert" = "red", "Heart" = "purple", "Mack" = "#155F83FF", "LEB" = "orange", "Marsh" = "magenta", "Hidden" = "pink", "Little_Lakes" = "gray", "Little_Fish" = "#4DAC26", "Matlock" = "#DFC27D", "No_Name" = "#D01C8B")) +
  theme_minimal()
```

some analysis by the date

```{r, all lakes on the same graph}
ggplot(summ_mos_counts, aes(x = date, y = effort_min, color = lake)) +
  geom_line() +
  labs(title = "Effort by Date", x = "Date", y = "Effort (minutes)") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, lakes on separate graphs}
ggplot(summ_mos_counts, aes(x = date, y = effort_min)) +
  geom_line() +
  labs(title = "Effort by Date", x = "Date", y = "Effort (minutes)") +
  theme_minimal() +
  facet_wrap(~lake)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, all lakes on same graph points not lines}
ggplot(summ_mos_counts, aes(x = date, y = effort_min, color = lake)) +
  geom_point() +
  labs(title = "Effort by Date", x = "Date", y = "Effort (minutes)") +
  theme_minimal() +
  facet_wrap(~lake)
```

